Unbalanced Mallows Models for Optimizing Expensive Black-Box Permutation Problems

### 2021-02-03 paper/main.tex
update heatmaps

\subsubsection{Datasets}
%
%For the LOP, we generated synthetic instances as follows. First, we create a sample $S$ of 200 permutations of size $n=20$ from a MM with $\sigma_0$ equal to the identity permutation and $\theta=-\log(\textit{Uniformity})$, where  $\textit{Uniformity}\in\{0.5,$ $0.7,$ $0.9\}$, i.e., from peaked distributions to almost uniform distributions. Then, we generate matrix $a$ as $a_{i,j} = \sum_{\sigma\in S}\sum_{i<j}\mathbb{I}\{\sigma(i)<\sigma(j)\}$, where ${\mathbb{I}\colon S \times S \to \{0,1\}}$ is the indicator function. In principle, these should be best-case instances for UMM since their fitness landscape closely matches the Mallows model.%\MANUEL{Ekhine: Could you revise this text? I wanted to explain why we use synthetic datasets.}
We consider real LOP datasets from the LOLIB
repository.\footnote{\url{http://grafo.etsii.urjc.es/optsicom/lolib/}} In
particular, we consider the following types: IO, Random instances of type A and
Random instances of type B of various sizes $n \in \{50, 100, 150, 200,
250\}$. These are harder problems than the synthetic instances, not only
because of the larger search space, but also because of the different instance
types.

In the case of the QAP and PFSP, we consider the same instances as
\citet{ZaeStoFriFisNauBar2014,ZaeStoBar2014:ppsn}, i.e., \texttt{nug12},
\texttt{nug30}, \texttt{tho30} and \texttt{kra32} for the QAP, with $n$ (the
permutation length) equal to the number in the instance name, and
\texttt{reC05}, \texttt{reC13} and \texttt{reC19} for the PFSP, with
$n \in \{20, 20, 30\}$, respectively. For reasons that will become obvious
later, we also consider QAP instances \texttt{kra30a} and \texttt{kra30b}.

@@

\begin{figure}[tb]
  \centering%
  \begin{minipage}{0.49\linewidth}\centering
    \texttt{N-t1d100.01} (LOP)\\[-0.2ex]
    \includegraphics[width=\textwidth]{../img/heatmap_N-t1d100_01}
  \end{minipage}
  \begin{minipage}{0.49\linewidth}\centering
    \texttt{N-t1d50.01} (LOP)\\[-0.2ex]
 \includegraphics[width=\textwidth]{../img/heatmap_N-t1d150_01}
\end{minipage}\\\vskip 1em
\begin{minipage}{0.49\linewidth}
  \centering
  \texttt{reC05} (PFSP)\\[-0.5ex]
  \includegraphics[width=\textwidth]{../img/heatmap_rec05}
\end{minipage}
% \begin{minipage}{0.49\linewidth}
%   \centering
%   \texttt{kra32} (QAP)\\[-0.5ex]
% \includegraphics[width=\textwidth]{../img/heatmap_kra32_dat}
% \end{minipage}
\caption{Mean fitness (over 10 runs) of different configurations of $r_1$ and $r_2$ (darker is better). Whitespace denotes missing data (runs that took longer than 6 hours).\label{fig:heatmaps}}
\end{figure}

>>>>

\subsubsection{Datasets}
%
%For the LOP, we generated synthetic instances as follows. First, we create a sample $S$ of 200 permutations of size $n=20$ from a MM with $\sigma_0$ equal to the identity permutation and $\theta=-\log(\textit{Uniformity})$, where  $\textit{Uniformity}\in\{0.5,$ $0.7,$ $0.9\}$, i.e., from peaked distributions to almost uniform distributions. Then, we generate matrix $a$ as $a_{i,j} = \sum_{\sigma\in S}\sum_{i<j}\mathbb{I}\{\sigma(i)<\sigma(j)\}$, where ${\mathbb{I}\colon S \times S \to \{0,1\}}$ is the indicator function. In principle, these should be best-case instances for UMM since their fitness landscape closely matches the Mallows model.%\MANUEL{Ekhine: Could you revise this text? I wanted to explain why we use synthetic datasets.}
We consider real LOP datasets from the LOLIB
repository.\footnote{\url{http://grafo.etsii.urjc.es/optsicom/lolib/}} In
particular, we have considered two instances of each class, focusing on relatively instances (the computation of CEGO grows rapidly with instance size):
\texttt{N-p40-01}, \texttt{N-p40-02}, \texttt{N-p50-01}, \texttt{N-p50-02} (RandB), 
lop/RandB/
lop/IO/N-t59d11xx
lop/IO/N-t59b11xx
lop/SGB/N-sgb75.02
lop/SGB/N-sgb75.01
RandB/
RandB/N-p50-02

: IO, Random instances of type A and
Random instances of type B of various sizes $n \in \{50, 100, 150, 200,
250\}$. These are harder problems than the synthetic instances, not only
because of the larger search space, but also because of the different instance
types.\MANUEL{Needs updating}

In the case of the PFSP, we consider the same instances as
\citet{ZaeStoBar2014:ppsn}, i.e., \texttt{reC05}, \texttt{reC13} and
\texttt{reC19} for the PFSP, with $n \in \{20, 20, 30\}$, respectively.

% In the case of the QAP and PFSP, we consider the same instances as
% \citet{ZaeStoFriFisNauBar2014,ZaeStoBar2014:ppsn}, i.e., \texttt{nug12},
% \texttt{nug30}, \texttt{tho30} and \texttt{kra32} for the QAP, with $n$ (the
% permutation length) equal to the number in the instance name, and
% \texttt{reC05}, \texttt{reC13} and \texttt{reC19} for the PFSP, with
% $n \in \{20, 20, 30\}$, respectively. For reasons that will become obvious
% later, we also consider QAP instances \texttt{kra30a} and \texttt{kra30b}.

@@

\begin{figure}[tb]
\centering%
  \begin{minipage}{0.49\linewidth}\centering
    \texttt{N-t59d11xx} (LOP)\\[-0.2ex]
    \includegraphics[width=\textwidth]{../img/heatmap_lop_IO_N-t59d11xx}
  \end{minipage}
  \begin{minipage}{0.49\linewidth}\centering
    \texttt{SGB_N-sgb75_02} (LOP)\\[-0.2ex]
 \includegraphics[width=\textwidth]{../img/heatmap_lop_SGB_N-sgb75_02}
\end{minipage}\\\vskip 1em
\begin{minipage}{0.49\linewidth}
  \centering
  \texttt{reC13} (PFSP)\\[-0.5ex]
  \includegraphics[width=\textwidth]{../img/heatmap_pfsp_rec13_txt}
\end{minipage}
\begin{minipage}{0.49\linewidth}
  \centering
  \texttt{reC19} (PFSP)\\[-0.5ex]
  \includegraphics[width=\textwidth]{../img/heatmap_pfsp_rec19_txt}
\end{minipage}
% \begin{minipage}{0.49\linewidth}
%   \centering
%   \texttt{kra32} (QAP)\\[-0.5ex]
% \includegraphics[width=\textwidth]{../img/heatmap_kra32_dat}
% \end{minipage}
\caption{Mean fitness (over 10 runs) of different configurations of $r_1$ and $r_2$ (darker is better).}\label{fig:heatmaps}}
\end{figure}
